---
title: "Bayesian data analysis - beta blocker cross-validation demo"
author: "Aki Vehtari"
date: "First version 2018-01-10. Last modified `r format(Sys.Date())`."
output:
  html_document:
    fig_caption: yes
    toc: TRUE
    toc_depth: 2
    number_sections: TRUE
    toc_float:
      smooth_scroll: FALSE
bibliography: modelsel.bib
csl: harvard-cite-them-right.csl
---

# Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE, message=FALSE, error=FALSE, warning=FALSE, comment=NA, out.width='95%')
```

**Load packages**
```{r}
library(tidyr)
library(rstanarm)
library(rstan)
options(mc.cores = parallel::detectCores())
library(loo)
library(shinystan)
library(ggplot2)
library(ggridges)
```

# Introduction

This notebook demonstrates a simple model we trust, that is, we assume
$M$-closed case [@Vehtari+Ojanen:2012]. In this case, cross-validation
is not needed and we we can get better accuracy using the explicit
model.

# Comparison of two groups with Binomial 

An experiment was performed to estimate the effect of beta-blockers on mortality of cardiac patients [the example is from @BDA3, Ch 3]. A group of patients were randomly assigned to treatment and control groups:

- out of 674 patients receiving the control, 39 died
- out of 680 receiving the treatment, 22 died

Data, where grp2 is a dummy variable that captures the difference of
the intercepts in the first and the second group.
```{r}
d_bin2 <- data.frame(N = c(674, 680), y = c(39,22), grp2 = c(0,1))
```

## Analysis of the observed data

To analyse whether the treatment is useful, we can use Binomial model for both groups and compute odds-ratio.

```{r,warning=FALSE,error=FALSE}
fit_bin2 <- stan_glm(y/N ~ grp2, family = binomial(), data = d_bin2,
                     weights = N, refresh=0)
```

Plot odds ratio
```{r}
samples_bin2 <- rstan::extract(fit_bin2$stanfit)
theta1 <- plogis(samples_bin2$alpha)
theta2 <- plogis(samples_bin2$alpha + samples_bin2$beta)
oddsratio <- (theta2/(1-theta2))/(theta1/(1-theta1))
ggplot() + geom_histogram(aes(oddsratio), bins = 50, fill = 'darkblue', color = 'black') +
  labs(y = '') + scale_y_continuous(breaks = NULL)
```

The probability that odds-ratio is less than 1:
```{r}
print(mean(oddsratio<1),2)
```

This posterior distribution of the odds-ratio (or some transformation
of it) is the simplest and the most accurate way to analyse the
effectivness of the treatment. In this case, there is high probability
that the treatment is effective.

## Simulation experiment

Simulation experiment `binom_odds_comparison.R` runs 100 simulations with simulated data with varying oddsratio (0.1,...,1.0) and computes for each run the probability that oddsratio<1. The follwing figures show the variation in the results.

Variation in probability that oddsratio<1 when true oddsratio is varied.
```{r}
load(file="binom_test_densities.RData")
ggplot(betaprobs_densities, aes(x = values, y = ind, height = scaled)) + 
  geom_density_ridges(stat = "identity", scale=0.6)
```

# Cross-validation

Sometimes it is better to focus on observable space (we can't observe
$\theta$ or odds-ratio directly, but we can observe $y$).  In
leave-one-out cross-validation, model is fitted $n$ times with each
observation left out at time in fitting and used to evaluate the
predictive performance. This corresponds to using the already seen
observations as pseudo Monte Carlo samples from the future data
distribution, with the leave-trick used to avoid double use of data.
With the often used log-score we get
$$\mathrm{LOO} = \frac{1}{n} \sum_{i=1}^n \log {p(y_i|x_i,D_{-i},M_k)}$$.
Cross-validation is useful in $M$-open case [@Vehtari+Ojanen:2012],
when we don't trust any model to present the future data well (they
might, but we just don't have enough confidenece to trust them to do
so).

Next we demonstrate one of the weakneses of cross-validation (same
holds for WAIC etc).

## Analysis of the observed data

To use leave-one-out where "one" refers to an individual patient, we
need to change the model formulation a bit. In the above model
formulation, the individual observations have been aggregated to group
observations and running `loo(fit_bin2)` would try to leave one group
completely. In case of having more groups, this could be what we want,
but in case of just two groups it is unlikely. Thus, in the following we 
switch a Bernoulli model with each individual as it's own observation.

```{r, warning=FALSE, error=FALSE}
d_bin2 <- data.frame(N = 674+680, y = c(rep(1,39), rep(0,674-39), rep(1,22), rep(0,680-22)), grp2 = c(rep(0, 674), rep(1, 680)))
fit_bin2 <- stan_glm(y ~ grp2, family = binomial(), data = d_bin2, seed=180202538, refresh=0)
```

We fit also a "null" model which doesn't use the group variable and thus has common parameter for both groups.'
```{r, warning=FALSE, error=FALSE}
fit_bin2null <- stan_glm(y ~ 1, family = binomial(), data = d_bin2, seed=180202538, refresh=0)
```

We can then use cross-validation to compare whether adding the
treatment variable improves predictive performance. We use fast Pareto smoothed importance sampling leave-one-out cross-validation [PSIS-LOO; @Vehtari+etal:PSIS-LOO:2017].
```{r}
(loo_bin2 <- loo(fit_bin2))
(loo_bin2null <- loo(fit_bin2null))
```
All Pareto $k<0.5$ and we can trust PSIS-LOO result [@Vehtari+etal:PSIS-LOO:2017; @Vehtari+etal:PSIS:2017].

Let's make pairwise comparison.
```{r}
compare(loo_bin2null, loo_bin2)
```
elpd_diff is small compared to se, and thus cross-validation is uncertain whether would be significant improvement in predictive performance! To put this in persepective, we have $N_1=674$ and $N_2=680$, and 5.8% and 3.2% deaths, and this is now too weak information for cross-validation.

## Simulation experiment

Simulation experiment `binom_odds_comparison.R` runs 100 simulations with simulated data with varying oddsratio (0.1,...,1.0) and computes LOO comparison for each run.

Variation in LOO comparison when true oddsratio is varied.
```{r}
ggplot(looprobs_densities, aes(x = values, y = ind, height = scaled)) + 
  geom_density_ridges(stat = "identity", scale=0.6)
```

We see that using the posterior distribution from the model is more
sensitive to detect the effect, but cross-validation will detect it
eventually too. The difference here comes that cross-validation
doesn't trust the model, compares the model predictions to the "future
data" using very weak assumption about the future. The weak assumption
about the future is also the cross-validation strength as we we'll see
in another notebook.

# Reference predictive approach

We can also do predictive performance estimates using stronger assumption about the future. A reference predictive estimate with log-score can be computed as
$$\mathrm{elpd}_{\mathrm{ref}}  = \int p(\tilde{y}|D,M_*) \log p(\tilde{y}|D,M_k) d\tilde{y}, $$
where $M_*$ is a reference model we trust. Using a reference model to assess the other models coreesponds to $M$-completed case [@Vehtari+Ojanen:2012], where the true model is replaced with a model we trust to be close enough to the true model.

## Simulation experiment

The next figure shows the results from the same simulation study using a reference predictive approach with the `fit_bin2` model used as the reference. 
```{r}
ggplot(refprobs_densities, aes(x = values, y = ind, height = scaled)) + 
  geom_density_ridges(stat = "identity", scale=0.6)
```

We can see better accuracy than for cross-validation. The similar improvement in the model selection performance is observed in projection predictive variable selection [@Piironen+Vehtari:2017a; @Piironen+etal:projpred:2018] implemented in [`projpred` package](https://cran.r-project.org/package=projpred).

<br />

# References {.unnumbered}

<div id="refs"></div>

# Licenses {.unnumbered}

* Code &copy; 2018, Aki Vehtari, licensed under BSD-3.
* Text &copy; 2018, Aki Vehtari, licensed under CC-BY-NC 4.0.
* Part of the code copied from [rstanarm_demo.Rmd](https://github.com/avehtari/BDA_R_demos/blob/master/demos_rstan/rstanarm_demo.Rmd) written by Aki Vehtari and Markus Paasiniemi

# Original Computing Environment {.unnumbered}

```{r}
sessionInfo()
```

<br />
